Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-06-10 FPT2014/bare_conf.tex
Davids comments on abstract intro and performance analysis

% author names and affiliations
\author{Authors removed for blind review\\ \\ \\}
%\author{\IEEEauthorblockN{Kan Shi, David Boland, and George A. Constantinides}
%\IEEEauthorblockA{Department of Electrical and Electronic Engineering\\
%Imperial College London\\London, UK\\
%\{k.shi11, david.boland03,g.constantinides\}@imperial.ac.uk}
%}

@@

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components are designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ area savings and $1.5\times$ speed-ups for multipliers. Furthermore in comparison to standard arithmetic primitives, the area overheads of online operators drop from $8.41\times$ to $1.88\times$ for adders, and $8.11\times$ to $0.54\times$ for multipliers using our approach.

@@

Despite its attractiveness, the use of online arithmetic on existing FPGAs is still limited due to a large area overhead. Given for standard arithmetic, it is difficult to efficiently map arithmetic circuits solely onto look-up-tables (LUTs), the fundamental building blocks of FPGAs. As a result, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce carry propagation delay~\cite{Virtex6}. Significant prior work explores methods that can effectively map circuits to the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or proposes alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2}. Hard DSP blocks are also included in modern architectures to improve the area and performance of multiplications and multiply accumulates~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, two major research directions have been investigated. First, although there has been previous research into FPGA implementations of online arithmetic, this work only focused on online adders; a brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. Second, there are existing approaches to build general purpose online adders specifically for FPGAs with 4-input LUTs and with two LUTs within a slice or a logic element (LE), such as the Xilinx Spartan~3 series \cite{XilinxSpartan} and the Altera Cyclone III \cite{AlteraCyclone}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and four LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}. Moreover, the move to the 6-LUT on the basic architecture opens new opportunities for efficient implementation.
%However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map the digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of a digit parallel online multiplier. The resource usage of both operators can be determined analytically. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. The area overhead drops from up to $8.41\times$ to $1.88\times$ using the proposed online adder in comparison to the standard ripple-carry adder. For online multipliers, the LUTs and slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. When compared to the multiplier with standard arithmetic, the direct implementation has an area overhead up to $8.11\times$. This value drops to $1.84\times$ when using the proposed online multiplier. Furthermore, using the proposed structure can even achieve area savings of up to $55\%$ for half precision results in comparison to the standard multiplier.

In summary, the main contributions of this paper are efficient methods to map online adders and multipliers to modern FPGAs. 
% by fully utilizing the fast carry logic for .

%\begin{enumerate}
%    \item An efficient approach of mapping online adders onto modern FPGAs;
%    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
%    \item Theoretical area models of the proposed architectures.
%\end{enumerate}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, as illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. $\delta$ is normally a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  \vspace{-4ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  \vspace{-1ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

@@

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes 3 inputs and generates 2 outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only 2 FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$ \cite{Ercegovac_OnlineMult}.
%
\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

\begin{eqnarray}\label{Eq:SelFunc_OM}
%\small
  sel(W_{[j]})=\begin{cases}
    1 & \text{ if } W_{[j]} \geqslant \frac{1}{2} \\
    0 & \text{ if } -\frac{1}{2}\leqslant W_{[j]}<\frac{1}{2} \\
    \overline{1} & \text{ if } W_{[j]}<-\frac{1}{2}
  \end{cases}
%\normalsize
\end{eqnarray}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan2.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The gray background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork2.pdf}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    %\vspace{-2ex}
\end{figure}

@@

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit2.pdf}
	\caption{Alternative structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. The dotted boxes outline the logic that can be mapped onto each LUT and the corresponding fast carry logic.}
	\vspace{-1ex}	
	\label{Fig:ModifiedOA}
\end{figure*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	%\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the area figures of the online adder in terms of the number of LUTs and slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OA\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OA\_slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. While in our experiments all designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1, our methods can also be applied on other FPGAs with same slice architecture such as the Xilinx 7 series FPGAs. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/Exp/Adder_Freq.eps}
	\caption{Rated frequencies of the RCA and the online adder with different implementation methods. The results are obtained from post place and route timing reports in ISE 14.1.}
	\label{Fig:AdderFreq}
	\vspace{-1ex}
\end{figure}

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Specifically, the frequency of \texttt{OA\_behv} is faster than \texttt{RCA} when $N\geqslant30$. In comparison, the proposed online adder is more efficient in performance, as it outperforms \texttt{RCA} for $N>16$. Furthermore, the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $1.20\times$ to $1.26\times$.

We also compare the resource usage of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig~\ref{Fig:AdderArea}. We can see that in comparison to the original \texttt{OA\_behv}, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in slices. This is because the dedicated FPGA resources are fully utilized. However, our design still costs more area than an \texttt{RCA}, with an area overhead of $80\%\sim98\%$ in terms of LUTs, and up to $88\%$ in terms of slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed.
%However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs3.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Usage of slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices3.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths.}
  \label{Fig:AdderArea}
  \vspace{-2ex}
\end{figure}

@@

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $z_{j}  ~~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$
  \end{algorithmic}
\end{algorithm}

@@

The blocks that generate partial products $Xy_j$ can be incorporated into the online adder for further area reduction. In the online adder structure as presented in Fig.~\ref{Fig:SliceNew}, we notice that only four inputs per LUT are used. This leaves at least two available inputs per LUT. Originally in order to generate one digit partial product, one digit of inputs $X$ and $Y$ are required respectively, as indicated in Fig.~\ref{Fig:PartialProduct}. Instead of using extra logic to generate $Xy_j$, it can be merged into the corresponding LUTs in the online adder by fully utilizing all six LUT inputs. Notice that this approach is only available for FPGAs with 6-LUTs.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.4\textwidth]{./Figures/PartialProduct.pdf}
	\vspace{-1ex}
	\caption{Left: Direct implementation with extra logic to generate partial products. Right: Combining the logic blocks that generate partial products into the Online Adder by fully utilizing all the inputs of the 6-LUTs. In this case, further area reduction can be achieved.}
	\label{Fig:PartialProduct}
	\vspace{-1ex}
\end{figure}

>>>>

% author names and affiliations
%\author{Authors removed for blind review\\ \\ \\}
%\author{\IEEEauthorblockN{Kan Shi, David Boland, and George A. Constantinides}
%\IEEEauthorblockA{Department of Electrical and Electronic Engineering\\
%Imperial College London\\London, UK\\
%\{k.shi11, david.boland03,g.constantinides\}@imperial.ac.uk}
%}

@@

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components were originally designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. More recently, research has shown that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve slice savings of over $67\%$ and $69\%$, and speed-ups of over $1.2\times$ and $1.5\times$ for adders and multipliers, respectively. Furthermore in comparison to standard arithmetic primitives, this translates to reduction of the area overheads of up to $8.41\times$ and $8.11\times$ for adders and multipliers using direct implementation, to $1.88\times$ and $0.54\times$ using our approach.

@@

Despite its attractiveness, the use of online arithmetic on existing FPGAs is still limited due to a large area overhead. For standard arithmetic, because it is difficult to efficiently map arithmetic circuits solely onto look-up-tables (LUTs), both main commercial FPGA vendors introduced dedicated carry logic to reduce carry propagation delay~\cite{Virtex6}. Significant prior work explores methods that can effectively map circuits to the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or proposes alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2}. Hard DSP blocks are also included in modern architectures to improve the area and performance of multiplications and multiply accumulates~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, two major research directions have been investigated. Firstly, there has been some research into FPGA implementations of online adders; a brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. Secondly, there are existing approaches to build general purpose online adders specifically for FPGAs with 4-input LUTs and with two LUTs within a slice or a logic element (LE), such as the Xilinx Spartan~3 series \cite{XilinxSpartan} and the Altera Cyclone III \cite{AlteraCyclone}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGAs, which typically have 6-input LUTs and four LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}. Moreover, the move to the 6-LUT on the basic architecture opens new opportunities for further optimization.
%However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic and ensures the available resources within a slice are fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient digit parallel online multiplier for FPGAs. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves  slice savings of over $67\%$ and  frequency speed-ups of over $1.2\times$ in comparison to a direct implementation. In comparison to the standard ripple-carry adder, the area overhead drops from up to $8.41\times$ to $1.88\times$ using the proposed online adder. For online multipliers, the slice utilization of the proposed architecture drops of over $69\%$, with a frequency speed-up over $1.5\times$ in comparison to a direct implementation. When compared to the multiplier with standard arithmetic, the direct implementation has an area overhead up to $8.11\times$, which drops to $1.84\times$ when using the proposed online multiplier.

Our approach has the additional advantage of computing a correctly rounded multiplier with a reduced output precision, whereas a standard multiplier has to generate full precision output. Using our method, a programmable number of output precision can be obtained, offering the ability to trade silicon area for precision. As an example, if we are only interested in the most significant half of the product, we can create a digit parallel online multiplier that saves upto $56\%$ slice usage in comparison to a standard multiplier with truncated output.

In summary, the main contributions of this paper are:
\begin{enumerate}
    \item Efficient methods to map online adders and multipliers to modern FPGAs;
    \item Demonstration of the performance improvements of online operators in terms of silicon area and operating frequency. 
\end{enumerate}

%In summary, the main contributions of this paper are efficient methods to map online adders and multipliers to modern FPGAs. The performance improvements in terms of silicon area and operating frequency with these new architectures mean that using online operators is much more attractive for an end user. The benefits of these operators .

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, as illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. $\delta$ is normally a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[htbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  \vspace{-4ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  \vspace{-1ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

@@

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$ \cite{Ercegovac_OnlineMult}.
%
\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}
%
\begin{eqnarray}\label{Eq:SelFunc_OM}
\small
  sel(W_{[j]})=\begin{cases}
    1 & \text{ if } W_{[j]} \geqslant \frac{1}{2} \\
    0 & \text{ if } -\frac{1}{2}\leqslant W_{[j]}<\frac{1}{2} \\
    \overline{1} & \text{ if } W_{[j]}<-\frac{1}{2}
  \end{cases}
\normalsize
\end{eqnarray}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan2.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The gray background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\vspace{-1ex}
	\label{Fig:PreviousWork}
\end{figure}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork2.pdf}
	\vspace{-1ex}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    \vspace{-1ex}
\end{figure}

@@

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit2.pdf}
	\vspace{-2ex}
	\caption{Alternative structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the two logic blocks (LBs) that can be mapped onto one slice. Right: one LB. The dotted boxes outline the logic that can be mapped onto each LUT and the corresponding fast carry logic.}
	\vspace{-2ex}	
	\label{Fig:ModifiedOA}
\end{figure*}

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.49\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\vspace{-5ex}
	\caption{Implementation of two logic blocks (LBs) in one FPGA slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	\vspace{-2ex}	
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the area figures of the online adder in terms of the number of LUTs and slices with respect to the operand word-lengths ($N$) can be calculated as given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.

\begin{eqnarray}\label{Eq:OAareaLUT}
\scriptsize
	OA\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OA\_slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\normalsize
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  We also compare these architectures against a ripple carry adder (\texttt{RCA}), which uses conventional arithmetic. While our experiments are targeting on Xilinx Virtex-6 FPGAs with speed grade -1, our methods are applicable to any FPGAs with the same slice architecture such as all the Xilinx 7 series FPGAs. Notice that in order to utilize the carry resources and avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/Exp/Adder_Freq.eps}
	\caption{Rated frequencies of the RCA and the online adder with different implementation methods. The results are obtained from post place and route timing reports in ISE 14.7.}
	\label{Fig:AdderFreq}
	\vspace{-1ex}
\end{figure}

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.7. It can be seen that both designs with online arithmetic achieve relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay of the online adder is theoretically independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. More specifically, the frequency of \texttt{OA\_behv} is faster than \texttt{RCA} when $N\geqslant30$, whereas the proposed online adder outperforms \texttt{RCA} for $N>16$. As a result, this makes the \texttt{OA\_new} more attractive for a wider range of applications. Furthermore, the speed-ups of \texttt{OA\_new} against \texttt{OA\_behv} are over $1.2\times$.

We also compare the resource usage of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig~\ref{Fig:AdderArea}. We can see that in comparison to the original \texttt{OA\_behv}, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in slices. This is because the dedicated FPGA resources are fully utilized. These area savings reduce the overhead of an \texttt{OA} over an \texttt{RCA} from upto $2.95\times$ to $1.98\times$ for LUTs, and from $8.41\times$ to $1.88\times$ for slices. It is unlikely that the overhead can be completely avoided when using an \texttt{OA} because of the redundant number system. Nevertheless, with the reduced overhead, using \texttt{OA} to take advantage of the graceful degradation when operating beyond the deterministic clocking region may become more attractive for an FPGA developer.

%However, our design still costs more area than an \texttt{RCA}, with an area overhead of $80\%\sim98\%$ in terms of LUTs, and up to $88\%$ in terms of slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed.
%However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs3.eps}
    \vspace{1ex}
  \end{minipage}%
  }\vspace{-1ex}
  \subfigure[Usage of slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices3.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths.}
  \label{Fig:AdderArea}
  \vspace{-2ex}
\end{figure}

@@

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $z_{j}  ~~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$					\vspace{.5ex}
  \end{algorithmic}
  \vspace{-1ex}
\end{algorithm}

@@

The blocks that generate partial products $Xy_j$ can be incorporated into the online adder for further area reduction. In the online adder structure as presented in Fig.~\ref{Fig:SliceNew}, we notice that only four inputs per LUT are used. This leaves at least two available inputs per LUT. Originally in order to generate one digit partial product, one digit of inputs $X$ and $Y$ are required respectively, as indicated in Fig.~\ref{Fig:PartialProduct}. Alternatively, instead of using extra logic to generate $Xy_j$, it can be merged into the corresponding LUTs in the online adder by fully utilizing all six LUT inputs, as shown on the right of Fig.\ref{Fig:PartialProduct}. Notice that this approach is only available for FPGAs with 6-LUTs.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.4\textwidth]{./Figures/PartialProduct.pdf}
	\vspace{-1ex}
	\caption{Left: Direct implementation with extra logic to generate partial products. Right: Combining the logic blocks that generate partial products into the Online Adder by fully utilizing all the inputs of the 6-LUTs. In this case, further area reduction can be achieved.}
	\label{Fig:PartialProduct}
	\vspace{-1ex}
\end{figure}
